Alex's Notes and Text Fragments for the JSS Manuscript

### 2024-02-28 paper/submission4/jstatsoft-sub4/article.Rmd

This toolbox, implemented as an open-source R package, builds upon and generalizes several existing toolboxes that have been developed in recent years for more specific settings (Table \ref{tab:sptcv-methods}).
The earliest and most comprehensive of these implementations is the \pkg{sperrorest} \proglang{R} package [@brenning2012], which provides an extensible framework and includes predefined resampling strategies based on geometric blocking, clustering, and buffering.
In contrast, packages \pkg{blockCV} and \pkg{ENMeval} were developed for block and buffer resampling with a focus on species distribution modeling [@blockCV; @rest2014; @muscarella2014].
However, neither of these have been integrated into established machine-learning frameworks such as \pkg{mlr3} [@mlr3] or \pkg{caret}/\pkg{tidymodels} [@kuhn2020], and all of them lack support for temporal prediction tasks.
The \pkg{CAST} package, in contrast, focuses on spatiotemporal prediction tasks and makes use of some functions of the \pkg{caret} framework [@cast; @meyer2018].
One limitation of all the packages which provide only a subset of the machine learning workflow, e.g., resampling methods, is the sole focus on model assessment.
Through the integration into the \pkg{mlr3} framework, \pkg{mlr3spatiotempcv} indirectly offers model selection capabilities, i.e., applying and evaluating various algorithms across different preprocessing setups (e.g., all the preprocessing ‘pipelining operators’ of \pkg{mlr3pipelines}) while being able to make use of parallel execution and enhanced logging abilities.
During the initial development of \pkg{mlr3spatiotempcv}, it has been the first toolbox aiming to integrate standalone spatiotemporal validation strategies into an existing machine-learning framework.
Meanwhile the \pkg{spatialsample} has undertaken similar efforts and added the ability to use spatial resampling methods for the \pkg{tidymodels} framework [@mahoney2023].
In addition parallel efforts to support spatial validation in the Python world have been made by the \pkg{spacv} and \pkg{spatial-kfold} packages, yet often with only a subset of available methods compared to what \pkg{mlr3spatiotempcv} offers [@ghariani2023].

@@

CV is also widely used to assess model performance when tuning hyperparameters of flexible machine-learning models or performing feature selection [@cai2018; @bischl2021].
The \pkg{mlr3} framework supports the use of CV for both approaches.
Using the CV methods introduced here, \pkg{mlr3} can therefore be used to optimized models to show an improved performance in specific spatial or spatiotemporal predictive setting [@schratz2019].
Such an optimization may, for example, result in a reduced maximum tree depth or increased minimum node size in the Ecuador case study, since these hyperparameter settings would result in a stronger generalization and reduced overfitting.
Usually both hyperparameter and feature selection (wrapper feature selection or filters) are combined within a single, nested optimization process [@schratz2021].

>>>>

This toolbox, implemented as an open-source R package, builds upon and generalizes several existing toolboxes that have been developed in recent years for more specific settings (Table \ref{tab:sptcv-methods}).
The earliest and most comprehensive of these implementations is the \pkg{sperrorest} \proglang{R} package [@brenning2012], which provides an extensible framework and includes predefined resampling strategies based on geometric blocking, clustering, and buffering.
In contrast, packages \pkg{blockCV} and \pkg{ENMeval} were developed for block and buffer resampling with a focus on species distribution modeling [@blockCV; @rest2014; @muscarella2014].
However, neither of these have been integrated into established machine-learning frameworks such as \pkg{mlr3} [@mlr3] or \pkg{caret}/\pkg{tidymodels} [@kuhn2020], and all of them lack support for temporal prediction tasks.
The \pkg{CAST} package, in contrast, focuses on spatiotemporal prediction tasks and makes use of some functions of the \pkg{caret} framework [@cast; @meyer2018].
One limitation of all the packages which provide only a subset of the machine learning workflow, e.g., resampling methods, is the sole focus on model assessment.
Through the integration into the \pkg{mlr3} framework, \pkg{mlr3spatiotempcv} offers model selection capabilities, i.e., the seamless evaluation and application of various algorithms across different preprocessing and optimization settings while being able to make use of parallel execution and enhanced logging abilities.
During the initial development of \pkg{mlr3spatiotempcv}, it has been the first toolbox aiming to integrate standalone spatiotemporal validation strategies into an existing machine-learning framework.
Meanwhile the \pkg{spatialsample} has undertaken similar efforts and added the ability to use spatial resampling methods for the \pkg{tidymodels} framework [@mahoney2023].
In addition parallel efforts to support spatial validation in the Python world have been made by the \pkg{spacv} and \pkg{spatial-kfold} packages, yet often with only a subset of available methods compared to what \pkg{mlr3spatiotempcv} offers [@ghariani2023].

@@

CV is also widely used to assess model performance when tuning hyperparameters or performing feature selection [@cai2018; @bischl2021].
The \pkg{mlr3} framework supports the use of CV for both approaches.
Using the CV methods introduced here, \pkg{mlr3} can therefore be used to optimize models to show an improved performance in specific spatial or spatiotemporal predictive settings [@schratz2019].
Such an optimization may, for example, result in a reduced maximum tree depth or increased minimum node size in the Ecuador case study, since these hyperparameter settings would result in a stronger generalization and reduced overfitting.
Usually both hyperparameter and feature selection (wrapper feature selection or filters) are combined within a single, nested optimization process [@schratz2021].
